Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-10-22 WSSSPE3/report/WSSSPE3_review.tex

WSSSPE3 broke free of the traditional Supercomputing Conference
environment this year, and in Turk's words, this spoke to the fact that scientific
software comes from many different types of inquiries, deployment, strategies
for maintenance, users; and ways of measuring the value of a piece of software.
It appeared to Turk that the supercomputing conferences generally adopt some
top-down approaches, whereas SciPy generally uses more bottom-up systems.
The essential messages
perceived were also often circumpolar: \katznote{this means ``situated around or inhabiting one of the earth's poles'' - not sure this is the right word} software getting harder in supercomputing
communities with exascale computing and optimization issues in mind, but software becoming better in the
SciPy group with emerging tools such as Jupiter and productivity packages for research workflow.
Admitting such comparisons are somewhat unfair generalizations, Turk reminded the
audience that the different approaches bring different types of ideas to the
table, but he welcomed WSSSPE3 being conducted outside existing preconceptions.

>>>>

WSSSPE3 broke free of the traditional Supercomputing Conference environment this
year, and in Turk's words, this spoke to the fact that scientific software comes
from many different types of inquiries, deployment, strategies for maintenance,
users; and ways of measuring the value of a piece of software. It appeared to
Turk that the supercomputing conferences generally adopt some top-down
approaches, whereas SciPy more often than not uses more bottom-up systems. The
essential messages perceived were also often bipolar: software getting harder in
supercomputing communities with exascale computing and optimization issues in
mind, but software becoming better in the SciPy group with emerging tools such
as Jupiter and productivity packages for research workflow. Admitting such
comparisons are somewhat unfair generalizations, Turk reminded the audience that
the different approaches bring different types of ideas to the table, but he
welcomed WSSSPE3 being conducted outside existing preconceptions.

Returning to the topic of his talk, Turk invited the audience to picture
scientific software as a flower on a landscape under the Sun, which may
represent a number of measurable factors such as number of
citations; growth of a community and number of contributors; amount of funding;
prestigious prizes awarded; stability of the community in terms of leadership
transitions, serving community needs, not breaking test suites, and performance
on new architectures. But all these metrics are strictly speaking \emph{proxies}
for the values and the impact scientific software bears. What we can measure
does not give us direct insight---it just gives us proxies of insight.

Turk then moved onto various different definitions of sustainability. His 
favorite one was keeping up with bug reports, in which case even if no new
features were added, the software remains sustainable. Another definition of
sustainability Turk mentioned was adding of new features, or maintaining the
software for a long period of time such as the cases of TeX or LaTeX with
community help. A notion Turk heard often at supercomputing conferences was that
sustainable software continued to work on new architectures. Yet another metric
was people continuing to be able to learn how to use and apply the software.
A funder Turk heard talked about sustainability as continuing to get funded. Turk also
recalled that Greg Wilson said in WSSSPE 1.1 that his view of sustainable software
was software that continued to give the same results over time. A last measure
of sustainability Turk presented was the ability to transition between different people
developing and using a piece of software.
